# Week 6 Exercises

yufan_yin_week6: 20.10. - 27.10.2020

Also see in the page to my course diary: <https://yufanyin.github.io/datavis-R/>

```{r include=F}
library(broom)
library(tidyverse)

```

## Exercise 1

The data frames df_w and df_f represent repeated measures data from 60 participants. Variables F1-F3 and W1-W3 are "sub-variables" that will be used to make two composite variables F_total and W_total, respectively. 

```{r include=F}
# simulating data
# This chunk run well after clicking 'run' but somehow had error when I knitted ('%>%' does not exist). So, I cannot submit html document.
set.seed(1)
df_f <- data.frame(id = rep(1:60, each = 2) %>% as.character,
                 session = rep(2:1, 60),
                 group = rep(1:2, each = 60),
                 F1 = sample.int(4, 120, replace = T) - 1,
                 F2 = sample.int(4, 120, replace = T) - 1,
                 F3 = (sample.int(4, 120, replace = T) - 1)) %>%
  mutate(F3 = F3*(session-1), F2 = F2*(group-1)*(2-session)) %>%
  mutate(group = factor(group))

df_w <- data.frame(id = rep(1:60, each = 2) %>% as.character,
                 session = rep(1:2, 60) %>% factor,
                 group = rep(1:2, each = 60) %>% factor,
                 W1 = sample.int(4, 120, replace = T) - 1,
                 W2 = sample.int(4, 120, replace = T) - 1,
                 W3 = sample.int(4, 120, replace = T) - 1)
```

### 1.1

Merge the two data frames together.

```{r}
df_f <- df_f  %>% 
  mutate(session = as.factor(session)) # many errors occurred when I tried to change the type of 'session' and 'group'. Q: I still did not understand why only factor works.

df <- full_join(df_f, df_w, by = c("id" = "id", "session" = "session", "group" = "group"), suffix = c("_f", "_w")) 

head(df)

```


### 1.2

Using the merged data frame, create the **composite variables F_total and W_total, which are the sums of F1-F3 and W1-W3, respectively** (i.e. their values can range from 0 to 9).

```{r}
df$F_total <- rowSums(df[, c('F1', 'F2', 'F3')])

df$W_total <- rowSums(df[, c('W1', 'W2', 'W3')]) 

# I searched all the material and did not find row or column sums were taught. Why it could be an exercise?

head(df)

```


## Exercise 2

### 2.1

Visualise the distributions of F_total and W_total for the two groups and measurement sessions (for example as boxplots). 

```{r}
df %>%
  ggplot(aes(session, F_total)) + 
  geom_boxplot() +
  facet_wrap(~group)

df %>%
  ggplot(aes(session, W_total)) + 
  geom_boxplot() +
  facet_wrap(~group)

```


```{r}
# try more
df %>%
  ggplot(aes(session, F_total)) + 
  geom_violin() +
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = .3, binwidth = .1) +
  facet_wrap(~group)

# Q: Is binwidth set without specific standard ('exploring multiple widths to find the best to illustrate the stories in your data')? I found the pots was too big when binwidth = 1 (according to data_wrangling_and_plotting_week3)?.

df %>%
  ggplot(aes(session, W_total)) + 
  geom_violin() +
  geom_jitter(alpha = .3) +
  facet_wrap(~group)

# Q: Is the distribution is the original without any calculating or rotation? If so, I prefer this plot to the next one.

# ['The jitter geom is a convenient shortcut for geom_point(position = "jitter"). It adds a small amount of random variation to the location of each point, and is a useful way of handling overplotting caused by discreteness in smaller datasets.']

df %>%
  ggplot(aes(session, W_total)) + 
  geom_violin() +
  geom_dotplot(binaxis = "y", stackdir = "center", alpha = .3, binwidth = .1) +
  facet_wrap(~group)

# ['stackdir: which direction to stack the dots. "up" (default), "down", "center", "centerwhole" (centered, but with dots aligned)']

```

### 2.2

Fit a linear regression model with F_total as the DV, and session and group as predictors.  

```{r}
# where is 'DV'?
# LM with an interaction effect
F_total.model.1 <- lm(F_total ~ session * group, data = df)

summary(F_total.model.1)

```

```
F_total_coef <- broom::tidy(F_total.model.1) %>%
  select(term, estimate) %>%
  mutate(estimate = round(estimate, 2)) %>% # round decimals for plot text
  spread(term, estimate) %>%
  rename(Intercept = `(Intercept)`,
         group_coef = group,
         session_coef = session,
         session_group_coef = `session:group`)

error: Can't rename columns that don't exist. x Column `group` doesn't exist.

```


```
# I failed to rename the columns above so this chunk had error. I have to delet {r} to retain the codes.

(F_total_plot <- broom::augment(F_total.model.1, se_fit = T) %>%
  ggplot(aes(session, F_total)) +
  geom_point(aes(color = group), alpha = .7) + 
  geom_line(aes(session, .fitted, color = group), size = 1) + 
  geom_ribbon(aes(ymin=.fitted-1.96*.se.fit, ymax=.fitted+1.96*.se.fit, fill = group), alpha=0.2) +
  theme_bw())

# plot annotations
F_total_plot +
  geom_point(aes(0, F_total_coef$Intercept)) + # mark the intercept point
  geom_text(aes(0.35, F_total_coef$Intercept, 
                label = paste("Intercept =", F_total_coef$Intercept)), vjust=-.9) +
  geom_text(aes(4.2, F_total_coef$Intercept + F_total_coef$session_coef * 4.2, # annotate session coefficient 
                label = paste("Slope =", F_total_coef$session_coef)),
            vjust = -.9) +
  geom_segment(aes(x = 1.3, y = F_total_coef$Intercept + F_total_coef$session_coef * 1.3, # draw arrow to mark gender coefficient
                   xend = 1.3, yend = F_total_coef$Intercept + F_total_coef$session_coef * 1.3 + F_total_coef$group_coef * 1), 
               arrow = arrow()) + 
  geom_text(aes(1.3, F_total_coef$Intercept + F_total_coef$session_coef * 1.3 + F_total_coef$gender_coef * 1, 
                label = paste("Female coef =", F_total_coef$group_coef)),
            vjust = 2, hjust = 1.1) +
   geom_segment(aes(x = 1.3, y = F_total_coef$Intercept + F_total_coefsession_coef * 1.3 + F_total_coef$group_coef * 1,
                    xend = 1.3, yend = F_total_coef$Intercept + F_total_coef$session_coef * 1.3 + GPA_coef$group_coef * 1 +  F_total_coef$session_group_coef * 1.3), 
               arrow = arrow()) +
   geom_text(aes(1.3, F_total_coef$Intercept + F_total_coef$session_coef * 1.3 + F_total_coef$group_coef * 1 + F_total_coef$session_group_coef * 1.3, 
                label = paste("Interaction coef =", F_total_coef$session_group_coef)),
            vjust = 2, hjust = 1.1) 

```

### 2.3

Look at the means of F_total by group and session. How are they linked to the linear regression model coefficients?

```{r}
grouped_F_total1 <- df %>%
  group_by(group)

grouped_F_total1 %>%
  summarise(F_total1 = mean(F_total))

grouped_F_total2 <- df %>%
  group_by(session)

grouped_F_total2 %>%
  summarise(F_total2 = mean(F_total))

```


## Exercise 3

Visualise the [anscombe dataset](https://rdrr.io/r/datasets/anscombe.html) using ggplot2.

```{r}
anscombe

anscombe$x_total <- rowSums(anscombe[, c('x1', 'x2', 'x3', 'x4')])

anscombe$y_total <- rowSums(anscombe[, c('y1', 'y2', 'y3', 'y4')]) 

head(anscombe)

```


```{r include=F}
# Exercise 4

# Prepare a **separate R Notebook/Markdown document**, which will be the first draft of your final assignment with your own data. In the draft, include the following:  

# 1. Outline the study design, your research question, DV(s), IV(s)  
# 2. Data wrangling: start from reading in the raw data and show all steps  
# 3. Plot your main result(s)    

# Even if you had already completed some of these steps before, please include all of them in your document. **NOTE: Return either a readable HTML document (.html or .nb.html), or an .Rmd file along with your data, to make it possible for us to review your work!** Make the document as professional-looking as possible (you can, of course, include your comments/questions in the draft). You will get feedback on the draft, based on which you can then make the final version. The final document should be a comprehensive report of your data wrangling process and results. 

# Sorry, recently I have other urgent task for my article and do not have enough time to do this part.

```
