# Week 4 Exercises

yufan_yin_week4: 6.10. - 12.10.2020

Also see in the page to my course diary: <https://yufanyin.github.io/datavis-R/>

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```

## Exercise 1 Histograms and density plots

### 1.1 Reading the data

Read the region_scores.csv data

```{r}
region_scores <- read.csv(file = "D:/Users/yinyf/datavis-R/week4/region_scores.csv", stringsAsFactors = TRUE)
region_scores <- region_scores %>%
  mutate(id = as.character(id),
         region = factor(region),
         education = factor(education, ordered = TRUE),
         gender = factor(gender))

glimpse(region_scores)

```

Cutting values (score) into intervals

to groups of width 10

```{r}
region_scores %>%
  mutate(score_group = cut_width(score, 10, boundary = 0)) %>% 
  count(score_group)

```


```{r}
region_scores <- region_scores %>%
  mutate(score_group = cut_width(score, 10, boundary = 0, 
                                 labels = c('-10','11-20','21-'))) 
```

```{r}
region_scores %>% 
  distinct(score_group)

```

Column `score_group` is not found.

```{r}
region_scores2 <- region_scores %>%
  group_by(education, score_group, .drop = FALSE) %>%
  summarise(mean_age = mean(age),
            sd_age = sd(age),
            n = n()) %>%
  ungroup()

region_scores2

```

### 1.2 Histograms

Create a figure that shows the distributions (density plots or histograms) of **age** and **score** in separate subplots (facets). What do you need to do first?

**Note:** I'm not sure the group varible to create subplots.

In the figure, set individual x-axis limits for age and score by modifying the `scales` parameter within `facet_wrap()`.

**Question:** What went wrong when I used facet_wrap() but saw the warning 'Layer 1 is missing `score_group(or other group variable)`' ? I met last week, too. I saved score_group.

```{r}
region_scores %>%
  ggplot(aes(age, fill = score_group)) + 
  geom_histogram(position = "identity", alpha = .5, binwidth = 1) 

```
(Try more as a reminder in future)

```{r}
region_scores %>%
  ggplot(aes(age, fill = gender)) + 
  geom_histogram(position = "identity", alpha = .5, binwidth = 1) 

```

```{r}
region_scores %>%
  ggplot(aes(score, fill = gender)) + 
  geom_histogram(position = "identity", alpha = .5, binwidth = 1) 

```

### 1.3 Density plots

**Note:** I do not understand the meaning of y-axis in such density plots.

```{r}
region_scores %>%
  ggplot(aes(age, fill = gender)) + 
  geom_density(alpha = .5) 

```

```{r}
region_scores %>%
  ggplot(aes(score, fill = gender)) + 
  geom_density(alpha = .5) 

```

## Exercise 2 Gather: wide-to-long, spread: long-to-wide and scatter plot

In this exercise, you will use the built-in iris dataset.  

```{r}
head(iris)
str(iris)

```

### 2.1 Make data into long format

Make the data into long format: gather all variables except species into new variables **var** (variable names) and **measure** (numerical values). You should end up with 600 rows and 3 columns (Species, var, and measure). Assign the result into `iris_long`.

```{r}
iris_long <- iris %>%
  gather(var, measure, -Species) 

```

```{r}
str(iris_long)

```

### 2.2 Spread: long-to-wide

In `iris_long`, separate **var** into two variables: **part** (Sepal/Petal values) and **dim** (Length/Width).  

Then, spread the measurement values to new columns that get their names from **dim**. You must create row numbers by dim group before doing this.  

You should now have 300 rows of variables Species, part, Length and Width (and row numbers). Assign the result into `iris_wide`.

**Note:** It was a bit complex than the example. I tried many times but failed. So I kept some of the codes in the following chunk.

```
iris_long %>%
  group_by(Species) %>%
  mutate(row = row_number()) %>%
  ungroup %>%
  spread(?, ?) %>%
  select(-row)

```

However,

```
Must extract column with a single valid subscript. x Subscript `var` has the wrong type `data.frame<Sepal.Width:double>`. i It must be numeric or character.

```

Or:

```{r}
iris_long %>%
  pivot_wider(names_from = c(var),
  values_from = measure) 

```
There is still error. 

### 2.3 Scatter plot

Using `iris_wide`, plot a scatter plot of length on the x-axis and width on the y-axis. Colour the points by part.

```
iris_wide %>%
  ggplot(aes(Length, Width), color = Species) + # x = length, y = width
  geom_point()
```


## Exercise 3 Read and summarize my own data

### 3.1 Reading my own data

Import your data into R.  Check that you have the correct number of rows and columns, column names are in place, the encoding of characters looks OK, etc.   

```{r}
learning2019_w4 <- read.csv(file = "D:/Users/yinyf/datavis-R/week0/learning2019_week4.csv", stringsAsFactors = TRUE)
```


### 3.2

Print the structure/glimpse/summary of the data. Outline briefly what kind of variables you have and if there are any missing or abnormal values. Make sure that each variable has the right class (numeric/character/factor etc).  

```{r}
learning_w4 <- learning2019_w4 %>%
  mutate(studylength = as.numeric(studylength),
         writingcourse = as.numeric(writingcourse))
str(learning_w4)

```


## Exercise 4 Counting observations by grouping variables

Pick a few (2-5) variables of interest from your data (ideally, both categorical and numerical).  

For **categorical variables**, count the observations in each category (or combination of categories). Are the frequencies balanced?  

```
learning19_w4 %>%
  count(cluster, gender) %>%
  arrange(desc(n)) %>%
  arrange(cluster)
```

Error: Must group by variables found in `.data`. * Column `cluster` is not found.
Neither is learning19_w4[1].
Well... I'm not very angry.


For **numerical variables**, compute some summary statistics (e.g. min, max, mean, median, SD) over the whole dataset or for subgroups. What can you say about the distributions of these variables, or possible group-wise differences?  

Overall:

```{r}
summary(learning_w4)

```

For subgroups:

**Note:" I do not believe the mean values of subgroups divided by gender or student status(Bechelor/Master) could be equal. What's wrong?

```{r}
grouped_df <- learning_w4 %>%
  group_by(studentstatus)

grouped_df %>%
  summarise(unref_mean = mean(learning_w4$unref), deep_mean = mean(learning_w4$deep), orga_mean = mean(learning_w4$deep))

```

We can see studylength (how many month students have been studied in the university) is a better grouping value than (numbers) of writingcourse. But …

Try cluster (student profile based on the combination of scores on 'unref', 'deep' and 'orga')

```{r}
learning_w4 %>%
  count(learning_w4[1])
```

```{r}
grouped_learning <- learning_w4 %>%
  group_by(learning_w4[1])

grouped_learning %>%
  summarise(unref_mean = mean(grouped_learning$unref), deep_mean = mean(grouped_learning$deep), orga_mean = mean(grouped_learning$orga))

# the results look strange but I do not know what went wrong

```

## Exercise 5 Visualise my own data

Describe if there's anything else you think should be done as "pre-processing" steps (e.g. recoding/grouping values, renaming variables, removing variables or mutating new ones, reshaping the data to long format, merging data frames together).

Do you have an idea of what kind of relationships in your data you would like to visualise and for which variables? For example, would you like to depict variable distributions, the structure of multilevel data, summary statistics (e.g. means), or include model fits or predictions?


### 5.1 Reading the data

**Structure of the data**

```{r, structure of the data}
learning2019 <- read.csv(file = "D:/Users/yinyf/datavis-R/week0/learning2019_w4.csv", stringsAsFactors = TRUE) 
learning19 <- learning2019[1:13]
str(learning19)
```

**The aim of the study** is to investigate the interrelationships between the approaches to learning and conceptions of academic writing among international university students. Altogether 218 international students of the university participated in the study in 2018 and 2019. Students were divided into homogeneous groups based on their Z scores on the three approaches to learning. Then we compare mean differences and ANOVA results between the profiles.

**The data 'learning2019'** consists of 218 observations and 17 variables. It contains their scores of approaches to learning (different ways that students process information: unreflective studying, deep approach to learning and organised studying), conceptions of academic writing (blocks, procrastination, perfectionism, innate ability, knowledge transforming and productivity), and some background information (categorical variables, eg:gender, age, faculty, student status and study length).

**The explanation of some columns** are as follows. Each of them was average value of 2-4 questions in 5-point Likert scale (1= totally disagree, 5 = fully agree).

- "unref": relying on memorisation in the learning process, lacking the reflective approach to studying and applying the fragmented knowledge base.

- "deep": comprehending the intentional content, using evidence and integrating with previous knowledge.

- "orga": time management, study organisation, effort management and concentration.

- "blocks": the inability to write productively whose reason is not intellectual capacity or literary skills.

- "procrastination": failing to start or postponing tasks like preparing for exams and doing homework.

- "perfectionism": setting overly high standards, pursuing flawlessness, and evaluating one’s behavior critically.

- "innateability": writing is a skill which "is determined at birth" or "cannot be taught or developed".

- "ktransforming": (knowledge transforming) using writing for developing knowledge and generating new ideas and in the reflective and dialectic processes.

- "productivity": (sense of productivity) part of self-efficacy in writing.

### 5.2 Exploring the data numerically and graphically

#### 5.2.1 Summaries of the variables

```{r}
summary(learning19)
```
#### 5.2.2 Relationships between the variables

Calculate and print the correlation matrix
```{r}
cor_matrix<-cor(learning19[2:10]) %>% round(digits = 2)
cor_matrix
```

Specialized according to the significant level and visualize the correlation matrix
p.mat <- cor.mtest(cor_matrix)$p

```{r}
library(corrplot)
p.mat <- cor.mtest(cor_matrix)$p
corrplot(cor_matrix, method="circle", type="upper",  tl.cex = 0.6, p.mat = p.mat, sig.level = 0.01, title="Correlations of learning19", mar=c(0,0,1,0))
```


#### 5.2.3 Creating an x-y scatter plot

```
learning19 %>%
  ggplot(aes(orga, procrastination, color = cluster)) + # x = orga, y = procrastination
  geom_point()

```


### 5.3 K-means clustering


#### 5.3.1 Calculate the distances

Euclidean distance matrix

```{r}
learning19_eu <- dist(learning19[2:4])
summary(learning19_eu)
```
#### 5.3.2 Determine the k

```{r}
set.seed(123)
k_max <- 5 # determine the number of clusters
twcss <- sapply(1:k_max, function(k){kmeans(learning19[2:4], k)$tot.withinss}) # calculate the total within sum of squares
qplot(x = 1:k_max, y = twcss, geom = 'line') # visualize the results
```

The twcss value decrease heavily from 2 - 5 clusters. The optimal number of clusters was 3.

#### 5.3.3 Perform k-means clustering

```{r}
learning19_km <- kmeans(learning19[2:10], centers = 3)
```

Plot the dataset with clusters

```{r}
pairs(learning19[2:10], col = learning19_km$cluster)
pairs(learning19[,2:4], col = learning19_km$cluster)
pairs(learning19[,5:10], col = learning19_km$cluster)
```

The optimal number of clusters was 3. We got the best overview with three clusters.


### 5.3.4 Perform k-means on the original data

```
library(devtools)
library(flipMultivariates)
learning19_scaled3 <- scale(learning19[2:4])
learning19_km3 <-kmeans(learning19_scaled3, centers = 3)
cluster <- learning19_km3$cluster
learning19_scaled3 <- data.frame(learning19_scaled3, cluster)
lda.fit_cluster <- lda(cluster ~ ., data = learning19_scaled3)
lda.fit_cluster
```

Warning in install.packages :
  package ‘flipMultivariates’ is not available
  
  
but I used to run it so I kept the codes.
  
```
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "orange", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}
```

```
classes3 <- as.numeric(learning19_scaled3$cluster)
plot(lda.fit_cluster, dimen = 2, col = classes3, pch = classes3, main = "LDA biplot using three clusters")
lda.arrows(lda.fit_cluster, myscale = 2)
```


### 5.3.5 3D plot

```
model_predictors <- dplyr::select(learning19_train, -deep2)
# check the dimensions
dim(model_predictors)
dim(lda.fit$scaling)
# matrix multiplication
matrix_product <- as.matrix(model_predictors) %*% lda.fit$scaling
matrix_product <- as.data.frame(matrix_product)
```

Next, install and access the plotly package. 

Create a 3D plot of the columns of the matrix product.

```
library(plotly)
plot_ly (x = matrix_product$LD1, y = matrix_product$LD2, z = matrix_product$LD3, type= 'scatter3d', mode='markers', color = learning19_train$deep2)
```

```{r, message=FALSE, warning=FALSE}
library(plot3D)

scatter3D(x = learning19$unref, y = learning19$deep, z = learning19$orga, col = NULL, 
          main = "learning19 data", xlab = "deep",
          ylab ="unref", zlab = "orga")
```


```{r, message=FALSE, warning=FALSE}
library(plotly)
plot_ly (x = learning19$unref, y = learning19$deep, z = learning19$orga, type= 'scatter3d', mode='markers', color = learning19$deep)
```
