# Week 2 Exercises

yufan_yin_week2: 23.9. - 28.9.2020

Also see in the page to my course diary: <https://yufanyin.github.io/datavis-R/>


## Exercise 1

### 1.1 Loading libraries and suppressing any output messages in the chunk settings

Create a new code chunk where you load the tidyverse package. In the chunk settings, suppress any output messages.

```{r,message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```

### 1.2 Reading the data

The tibble *df* has 60 observations (rows) of variables (columns) group, gender, age, score1 and score2 (continuous scores from two tests). Each row represents one participant.  
```{r include = F}
# this chunk generates the data, and you can ignore it
set.seed(1)
df  <- tibble(group = sample(c(rep(1:2, each=30))),
              gender = sample(c(rep(c('M', 'F'), 28), rep(999,4))),
              score1 = rnorm(60, 18, 2),
              score2 = as.character(runif(60, 0.7, 1.1)*score1))
  
```

```{r}
df
```


There is something to fix in three of the variables. Explore the data and describe what needs to be corrected.  

**Hint**: You can use e.g. str(), distinct(), and summary() to explore the data.

```{r}
str(df)
summary(df)
distinct(df)
```

The dataset *df* consists of 60 observations and 5 variables.It contains the membership of group, gender, age, score1, score2.


## Exercise 2

### 2.1 Tidying data

Make the corrections you described above.

```{r}

df <- df %>%
  mutate(gender = na_if(gender, 999)) # recode 999 to NA (missing)
  df$score2 <- as.numeric(df$score2) # convert a character vector to a numeric vector
  
```

### 2.2 Counting observations by grouping variables

Count observations by group and gender. Arrange by the number of observations (ascending).
```{r}

df %>%
  count(group, gender) %>% # count() is a combination of group_by() and tally()
  arrange(desc(n)) %>% # OR: "%>% floor()"?
  arrange(group)

```


## Exercise 3

### 3.1 Creating a new variable: the difference between scores

Create a new variable, **score_diff**, that contains the difference between score1 and score2.

```{r}

df$score_diff <- df$score1 - df$score2

```

### 3.2 Computing the means: using summarise() to take multiple variables in one go

Compute the means of score1, score2, and score_diff.

**Hint**: Like mutate(), summarise() can take multiple variables in one go.

```{r}
df %>%
  summarise(score1_mean = mean(df$score1), score2_mean = mean(df$score2), score_diff_mean = mean(df$score_diff))

```

### 3.3 Computing the means by grouping variable

Compute the means of score1, score2, and score_diff **by gender**.

```{r, warning=FALSE, message=FALSE}
grouped_df <- df %>%
  group_by(gender)

grouped_df %>%
  summarise(score1_mean = mean(df$score1), score2_mean = mean(df$score2), score_diff_mean = mean(df$score_diff))

  # the results look strange but I do not know what went wrong
```


## Exercise 4

### 4.1 Creating an x-y scatter plot

Using ggplot2, create a scatter plot with score1 on the x-axis and score2 on the y-axis. 

```{r}
df %>%
  ggplot(aes(score1, score2)) + # x = score1, y = Sscore2
  geom_point()

```

### 4.2 Setting colour based on grouping variable, figure width and height

Continuing with the previous plot, colour the points based on gender. 

Set the output figure width to 10 and height to 6.


```{r, fig1, fig.height = 6, fig.width = 10}
df %>%
  ggplot(aes(score1, score2, color = gender)) + # x = score1, y = score2
  geom_point()

```


## Exercise 5

**Note:** I did this part in another rmd file named 'index'.

see: <https://github.com/yufanyin/datavis-R/blob/master/index.Rmd>

### 5.1 Metadata section

Add the author (your name) and date into the metadata section. Create a table of contents.


### 5.2 Knitting

Knit your document to HTML by changing `html_notebook` to `html_document` in the metadata, and pressing Knit.   

See the results in my course diary: <https://yufanyin.github.io/datavis-R/>

