# Final Assignment


```{r include=F}
library(broom)
library(tidyverse)

```

Prepare a **separate R Notebook/Markdown document**, which will be the first draft of your final assignment with your own data. In the draft, include the following:  

1. Outline the study design, your research question, DV(s), IV(s)  
2. Data wrangling: start from reading in the raw data and show all steps  
3. Plot your main result(s)    

Even if you had already completed some of these steps before, please include all of them in your document. **NOTE: Return either a readable HTML document (.html or .nb.html), or an .Rmd file along with your data, to make it possible for us to review your work!** Make the document as professional-looking as possible (you can, of course, include your comments/questions in the draft). You will get feedback on the draft, based on which you can then make the final version. The final document should be a comprehensive report of your data wrangling process and results. 


## 7.1 Reading the data

**Structure of the data**

```{r, reading the data}
learning2019 <- read.csv(file = "D:/Users/yinyf/datavis-R/week0/learning2019_w4.csv", stringsAsFactors = TRUE) 
learning19 <- learning2019[1:13]
str(learning19)
```

```
**The aim of the study** is to investigate the interrelationships between the approaches to learning and conceptions of academic writing among international university students. The research questions are as follows:

1) What kinds of conceptions of academic writing (blocks, procrastination, perfectionism, innate ability, knowledge transforming and productivity) the international students have?

2) What kinds of approaches to learning (deep approach to learning, unreflective studying and organised studying) the international students apply in their studies and what learning profiles can be identified?

3) What are the differences in conceptions of academic writing between the learning profiles?

Altogether 218 international students of the university participated in the study in 2018 and 2019. Students were divided into homogeneous groups based on their Z scores on the three approaches to learning. Then we compare mean differences and ANOVA results between the profiles.

```

```
The data were collected with Writing Process Questionnaire (Lonka, 2003; Lonka et al., 2014) and HowULearn Questionnaire (Parpala & Lindblom-Ylänne, 2012; Hailikari & Parpala, 2014). Both of them are the 5-point Likert scale and have been validated in Finnish and other contexts. 

**The data 'learning2019'** consists of 218 observations and 17 variables. It contains their scores of approaches to learning (different ways that students process information: unreflective studying, deep approach to learning and organised studying), conceptions of academic writing (blocks, procrastination, perfectionism, innate ability, knowledge transforming and productivity), and some background information (categorical variables, eg:gender, age, faculty, student status and study length).

**The explanation of some columns** are as follows. Each of them was average value of 2-4 questions in 5-point Likert scale (1= totally disagree, 5 = fully agree; hence it means nothing if I calculate the sum of A1, A2 and A3 or AW1 - AW6).

- "cluster": the membership after doing k-means clustering analysis. 1) Reflective and organised students (N=97, 45.1%), 2) Reflective and unorganised students (N=78, 36.3%), and 3) Unreflective and unorganised students (N=40, 18.6%)

- "unref": relying on memorisation in the learning process, lacking the reflective approach to studying and applying the fragmented knowledge base.

- "deep": comprehending the intentional content, using evidence and integrating with previous knowledge.

- "orga": time management, study organisation, effort management and concentration.

- "blocks": the inability to write productively whose reason is not intellectual capacity or literary skills.

- "procrastination": failing to start or postponing tasks like preparing for exams and doing homework.

- "perfectionism": setting overly high standards, pursuing flawlessness, and evaluating one’s behavior critically.

- "innateability": writing is a skill which "is determined at birth" or "cannot be taught or developed".

- "ktransforming": (knowledge transforming) using writing for developing knowledge and generating new ideas and in the reflective and dialectic processes.

- "productivity": (sense of productivity) part of self-efficacy in writing.

```

## 7.2 Relationships between the variables: Correlation matrix

Calculate and print the correlation matrix
```{r}
cor_matrix<-cor(learning19[2:10]) %>% round(digits = 2)
cor_matrix
```

Specialized according to the significant level and visualize the correlation matrix
p.mat <- cor.mtest(cor_matrix)$p

```{r, message=FALSE}
library(corrplot)
p.mat <- cor.mtest(cor_matrix)$p
corrplot(cor_matrix, method="circle", type="upper",  tl.cex = 0.6, p.mat = p.mat, sig.level = 0.01, title="Correlations of learning19", mar=c(0,0,1,0))
```


## 7.3 Learning profiles (based on the combination of approaches to learning)

### 7.3.1 Means of approaches to learning

```{r, add descriptive labels, message=FALSE}
learning19$锘縞luster <- factor(learning19$锘縞luster, levels = c("1", "2" , "3"),
                 labels = c("Reflective and organised students", "Reflective and unorganised students", "Unreflective and unorganised students"))

learning19$gender <- factor(learning19$gender, levels = c("1", "2"),
                 labels = c("Male", "Female"))

```

```{r, means of approaches to learning, message=FALSE}
grouped_df1 <- learning19 %>%
  group_by(锘縞luster)

grouped_df1 %>%
  summarise(unref_mean = mean(unref), deep_mean = mean(deep), orga_mean = mean(orga))

```

### 7.3.2 Boxplot: means of approaches to learning

```{r, message=FALSE, warning=FALSE, fig.width= 10}
learning19 %>%
  ggplot(aes(锘縞luster, unref)) + 
  geom_boxplot() +
  facet_wrap(~gender) + 
  labs(x = "Cluster", y = "Unreflective studying", title = "Means of unreflective studying") + 
  theme(axis.text.x = element_text(angle = 15))

learning19 %>%
  ggplot(aes(锘縞luster, deep)) + 
  geom_boxplot() +
  facet_wrap(~gender) + 
  labs(x = "Cluster", y = "Deep approach to learning", title = "Means of deep approach to learning") + 
  theme(axis.text.x = element_text(angle = 15))

learning19 %>%
  ggplot(aes(锘縞luster, orga)) + 
  geom_boxplot() +
  facet_wrap(~gender) + 
  labs(x = "Cluster", y = "Organised studying", title = "Means of organised studying") + 
  theme(axis.text.x = element_text(angle = 15))

# the descriptive labels of cluster were too long to display. I tried some codes searching online and finally they worked.

```

## 7.4 Differences in conceptions of academic writing between the learning profiles

### 7.4.1 Means of conceptions of academic writing

```{r, means of conceptions of academic writing, message=FALSE}
grouped_df2 <- learning19 %>%
  group_by(锘縞luster)

grouped_df2 %>%
  summarise(blocks_mean = mean(blocks), proc_mean = mean(procrastination), perf_mean = mean(perfectionism), inab_mean = mean(innateability), ktrans_mean = mean(ktransforming), produ_mean = mean(productivity))

```


### 7.4.2 X-y scatter plot: orga-procrastination;orga-productivity

```{r, x-y scatter plot 1, message=FALSE, fig.width= 10}
learning19 %>%
  ggplot(aes(orga, procrastination, color = 锘縞luster)) + # x = orga, y = procrastination
  geom_jitter(alpha = .5) +
  labs(x = "Organised studying", y = "Procrastination", title = "Organised studying - procrastination in different learning profiles") + 
  geom_point()

```


```{r, x-y scatter plot 2, message=FALSE, fig.width= 10}
learning19 %>%
  ggplot(aes(orga, productivity, color = 锘縞luster)) + # x = orga, y = productivity
  geom_jitter(alpha = .5) +
  labs(x = "Organised studying", y = "Productivity", title = "Organised studying - productivity in different learning profiles") + 
  geom_point()

```


### 7.4.3 Bar plots: cluster, blocks

```{r, message=FALSE, fig.width= 10}
learning19 %>% 
  ggplot(aes(锘縞luster, blocks, fill = gender)) + 
  geom_col(position = "dodge") + 
  labs(x = "Cluster", y = "Blocks", title = "Blocks in different gender and learning profiles") + 
  theme(axis.text.x = element_text(angle = 15)) +
  geom_smooth(method = "lm")

```

### 7.4.4 Bar plots: deep, perfectionism

```{r, message=FALSE, fig.width= 10}
grouped_df3 <- learning19 %>%
  group_by(deep)

grouped_df3 %>%
  summarise(perfectionism_mean2 = mean(perfectionism))

learning19 %>% 
  ggplot(aes(deep, perfectionism, fill = 锘縞luster)) + 
  geom_col(position = "dodge", alpha = .5) + 
  labs(x = "Deep approach to learning", y = "Perfectionism", title = "Deep approach to learning and perfectionism in different learning profiles")

# the values of y-axis are the means of perfectionism per each level of deep approach. However, I was unable to use the calculating result or summary_stat().

```

### 7.4.5 One-way ANOVA test: blocks

one of the findings is that students applying deep approach to learning experienced less blocks and perfectionism and tend not to regard academic writing as an innate ability.

Take deep~blocks as an example:

```{r}
aov_blocks <- aov(blocks ~ 锘縞luster, data = learning19)
summary(aov_blocks)

```

### 7.4.6 Simple regression with lm(): blocks ~ deep

```{r}
lm_blocks_deep <- lm(blocks~deep, data = learning19)
summary(lm_blocks_deep)

```

There is a statistical relationship between deep approach to learning and blocks (p:6.72e-05). Then draw a scatter plot, fit a linear model and print out the summary.

```{r}
qplot(deep, blocks, data = learning19) + geom_smooth(method = "lm")
model_blocks_deep <- lm(blocks ~ deep, data = learning19)
model_blocks_deep

```

### 7.4.7 Multiple regression: blocks ~ deep + orga

This multiple regression is to test whether organised studying has an influence on blocks.

```{r}
model_blocks_deep_orga <- lm(blocks ~ deep + orga, data = learning19)
summary(model_blocks_deep_orga)

```
The p-value (0.03) shows that the significance of the influence was statistical. The multiple R-squared (0.09) is a bit higher than that in regression model of blocks ~ deep (0.07). That means a little higher correlations if organised studying is taken into account.

I used to conduct modelling of approaches to learning and other variables in the data collected with the same questionnaire. The multiple R-squared was 0.20 so I do not know whether 0.09 is insufficient.

```
Thank you for your kind feedback these weeks.

```
